Reproducible Research: Peer Assessment 1 
================================================================================================

Loading and preprocessing the data

```{r}

Sys.setlocale("LC_TIME", "English")
file <- "activity.zip"
df <- read.csv(unz(file,  "activity.csv"))
df$date <- as.Date(df$date, format = "%Y-%m-%d")

```

### What is mean total number of steps taken per day?

```{r}

sumSteps <- aggregate(steps ~ as.Date(date), data = df, FUN=sum)
hist(sumSteps$steps, xlab = "total num of steps per day", breaks=30, 
     main = "Histogram of the total number of steps taken per day")
abline(v = mean(sumSteps$steps, na.rm = TRUE), col = "red", lwd = 2 )
abline(v = median(sumSteps$steps, na.rm = TRUE), col = "blue", lwd = 1 )
meanSteps <- mean(sumSteps$steps, na.rm = TRUE)

```

The mean total number of steps taken per day is `r sprintf("%.2f", meanSteps)`

### What is the average daily activity pattern?

```{r}

stepsInterval_mean <- aggregate(steps ~ interval, data = df, FUN=mean)
names(stepsInterval_mean) <- c("interval", "steps_mean")
plot(stepsInterval_mean, type = "l", main="The average number of steps")
maxStepsIntrval <- stepsInterval_mean$interval[which.max(stepsInterval_mean$steps_mean)]

```

The 5-minute interval that contains the maximum number of steps is : `r maxStepsIntrval`

### Imputing missing values

```{r}

missing_values <- sum(is.na(df))
dfNAajusted <- merge(df, stepsInterval_mean, by="interval")
dfNAajusted$fixSteps <- ifelse(!is.na(dfNAajusted$steps), dfNAajusted$steps, dfNAajusted$steps_mean)
sumSteps <- aggregate(steps ~ as.Date(date), data = dfNAajusted, FUN=sum)
hist(sumSteps$steps, xlab = "total num of steps per day", breaks=30,
     main = "Histogram of the total number of steps taken per day")
abline(v = mean(sumSteps$steps, na.rm = TRUE), col = "red", lwd = 2 )
abline(v = median(sumSteps$steps, na.rm = TRUE), col = "blue", lwd = 1 )
meanSteps <- mean(sumSteps$steps, na.rm = TRUE)
```

The total number of rows with missing values in the dataset is : `r missing_values`


New mean total number of steps taken per day is `r sprintf("%.2f", meanSteps)`

### Are there differences in activity patterns between weekdays and weekends?

```{r}

df$dayType <- weekdays(df$date)
df$dayType <- ifelse (df$dayType == "Saturday" | df$dayType == "Sunday", c("weekend"), c("weekday"))
stepsInt <- aggregate(steps ~ interval + dayType, data=df, FUN=mean)
names(stepsInt) <- c("interval", "dayType", "steps")

library(lattice)
xyplot(steps ~ interval | dayType, stepsInt, type = "l", layout = c(1, 2))

```
